---
title: "Homework 1"
author: "Beau Harrison"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

http://www.trfetzer.com/wp-content/uploads/Homework1.pdf

Parameters:

pid (debate id), url, sequence within debate, debate, date, speaker, what was spoken (each fragment separately)

### Libraries

```{r}
library(xml2) # html_nodes
library(rvest) # %>% notation
library(data.table)
```

### Get links within desired datespan

This block finds link nodes based on date

I'm using REGEX to find those debates from this election but I can remove this filter to return all elections

I'm not removing this filter because populating my data.table is extremely slow already

I find all rows with the class of 'docdate' and validate their text value against a regex date outputting only the rows I'm interested in

```{r}
raw.html <- read_html('http://www.presidency.ucsb.edu/debates.php')
html.rows <-  raw.html %>%
              xml_nodes('tr')
html.docdates <-  html.rows %>%
                  xml_nodes('.docdate')
validDateRows <- grep('(January|February|March|April|May|June|July|August|September|October|November|December) [0-9]*(st|nd|rd|th)?, (2015|2016)', html.docdates)
linkRows <- sapply(html.docdates[validDateRows], function(x) max(grep(x, html.rows)))
```

### Get attributes from each link

#### url

From the rows returned above find all anchor tags and return the value of the href property, url

```{r}
urls <- sapply(html.rows[linkRows], function(x) x %>% xml_nodes('a') %>% xml_attr('href'))
```

#### pid

Use a string split to parse the pid from the url above

```{r}
getPid <- function(url) return(unlist(strsplit(url, '='))[2])
```

#### date

Get the text from the docdates to use as the dates for each debate

```{r}
getDate <- function(htmlPage) return(htmlPage %>% html_node('.docdate') %>% xml_text())
```

#### debate

Fortuantely there was a unique class name for the title of the debates

I use the paperstitle class to get the debate name for each debate

```{r}
getDebate <- function(htmlPage) return(htmlPage %>% html_node('.paperstitle') %>% xml_text())
```

#### speaker

To find each speaker I find each paragraph and look for a bold tag inside the paragraph

If there is not a bold tag I ignore the line returned

I also do some clean up by parsing the colon out of each name

```{r}
getSpeakers <- function(htmlPage) { #xml_child seems to be faster than html_node
  ps <- htmlPage %>% xml_nodes('p')
  bs <- unlist(lapply(ps[2:length(ps)], function(p) p %>% xml_child('b') %>% xml_text()))
  newSpeakerLines <- which(!is.na(bs))
  speakers <- unlist(lapply(newSpeakerLines, function(lineIndex) gsub(':','',bs[lineIndex])))
  return(speakers)
}
```

#### text

Here I use a similar strategy to the names but I'm parsing out the names

I also use the lines that I found the names on to combine p tags until I find another line with a name and create a new text block

```{r}
getTexts <- function(htmlPage) { #xml_child seems to be faster than html_node
  ps <- htmlPage %>% xml_nodes('p')
  bs <- unlist(lapply(ps[2:length(ps)], function(p) p %>% xml_child('b') %>% xml_text()))
  newSpeakerLines <- which(!is.na(bs))
  psTextWNames <- ps %>% xml_text()
  psText <- gsub('^[A-z]+:', '', psTextWNames[2:length(psTextWNames)]) # parse out names
  newSpeakerLines <- c(newSpeakerLines, length(psText)) # append index for last line of text
  speakerText <- unlist(lapply(1:(length(newSpeakerLines)-1), function(lineNumber) paste(psText[newSpeakerLines[lineNumber]:(newSpeakerLines[lineNumber+1]-1)], sep = ' ', collapse = '')))
  return(speakerText)
}
```

#### combined

Parameters:

pid (debate id), url, sequence within debate, debate, date, speaker, what was spoken (each fragment separately)

This was very slow with data.table even after unlisting all my lists

```{r}
pidList <- integer()
urlList <- character()
seqList <- integer()
debateList <- character()
dateList <- character()
speakerList <- character()
textList <- character()

for(i in 1L:length(urls)) {
  htmlPage <- read_html(urls[i])
  pid <- getPid(urls[i])
  speakers <- getSpeakers(htmlPage)
  speakerList <- c(speakerList, speakers)
  texts <- getTexts(htmlPage)
  textList <- c(textList, texts)
  debate <- getDebate(htmlPage)
  date <- getDate(htmlPage)
  
  for(j in 1L:length(speakers)) {
    pidList <- c(pidList, pid)
    urlList <- c(urlList, urls[i])
    seqList <- c(seqList, j)
    debateList <- c(debateList, debate)
    dateList <- c(dateList, date)
  }
}

final <- data.table(pid=pidList,url=urlList,seq=seqList,debate=debateList,date=dateList,speaker=speakerList,text=textList)
```

### Results

#### Number of Debates

```{r}
sum(!duplicated(final$debate))
```

#### Number of Speakers

```{r}
sum(!duplicated(final$speaker))
```

#### head() of data.frame or data.table objects

```{r}
head(final)
```
